### 2011-02-09 latex/paper.tex
finalna verzia

\documentclass{ifacconf}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{graphicx}
\usepackage{natbib}

\bibpunct{(}{)}{;}{a}{,}{}

\usepackage{color}
\usepackage{subfig}

%nastavenie vypisov kodu
\usepackage{listings}
\usepackage{caption}

\captionsetup[lstlisting]{singlelinecheck=false}

\lstset{
language=Python, 
basicstyle=\footnotesize,
numbers=left,
numberstyle=\tiny,
xleftmargin= 12pt,
abovecaptionskip=10pt,
stepnumber=1,
numbersep=5pt,
frame=none,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=true,
escapeinside={\%}{)}
}

\begin{document}
\begin{frontmatter}
\title{Simulation of 2D physics of objects captured by web camera using OpenCV
and Box2D}
\author[Bratislava]{Michal Sedlák}
\address[Bratislava]{Faculty of Electrical Engineering and Information
Technology, Slovak University of Technology, Ilkovičova 3, 812 19 Bratislava,
Slovakia
\\
(e-mail: michal.sedlak@stuba.sk)}
\begin{abstract}
The paper presents one approach to simulation of physics applied on objects
captured by web camera. Introduced approach utilise OpenCV library for image
capturing and contour detection. Objects detected by OpenCV are reconstructed
from its outlines in Box2D environment so the physics can be applied to it.
Because of restrictions of Box2D we needed to do approximation and scaling of
outlines and tessellation of objects with Delaunay triangulation algorithm.
\end{abstract}
\begin{keyword}
OpenCV, Python, Box2D, physics, tessellation
\end{keyword}
\end{frontmatter}

\section{Introduction}
This paper describes applying of Newtonian physics to objects
recognized in image captured from camera. Simulation of physics is used in many
modern applications. You can find it in implementations used by game
engines, there are as well more complex implementation used in 3D drawing and
animation programs or exact and precise simulation in CAE and CAD programs.
Paper describes process of animation of objects, from a capturing
phase, over detection of the object outlines and interpretation of objects in 
physical engine, to animation of such objects. This approach can be applied in
education of physics at elementary schools, with interactive blackboards, or in
computer games.

\section{Object detection and Open Computer Vision library}
To apply a physics to hand drawn objects we need to identify and isolate
objects from image. We have used a web camera as a source and Open Computer
Vision (OpenCV) library as processing tool of the images.

\subsection{OpenCV}
In regards the book of \cite{OpenCV} OpenCV  is a library for open source
programming functions for real time computer vision, with more than five hundred
optimized algorithms. It can be used with C++, C and Python. We chose Python
version, which is optimized Python wrapper to C++ functions.

In the beginning we have to capture image to work with. OpenCV library has
implemented methods for image capture from camera. Simple image capture is shown
in Listing~\ref{Capture}.

\begin{lstlisting}[caption=Query image frame from web camera, label=Capture]
self.camera = cv.CaptureFromCAM(-1) %\label{Capture:init})
self.image = cv.QueryFrame(self.camera) %\label{Capture:query})
self.DetectOutline(self.image)
\end{lstlisting}

In line \ref{Capture:init} of Listing~\ref{Capture} we initialize our web
camera. In variable \lstinline{camera} is allocated and initialized object that
can query web camera for new image. Then as we see in Listing~\ref{Capture}
line~\ref{Capture:query} we can get the image from camera and store it in the
variable named \lstinline{image}. Captured image is shown in
Fig.~\ref{fig:captured}.

\begin{figure}[h]
\center
\includegraphics[width=0.8\columnwidth]{images/1test.png}
\caption{Image captured form camera}
\label{fig:captured}
\end{figure}

Now when we have image data stored in the variable, we can process data to find
outlines.

\begin{lstlisting}[caption=Outline detection, label=Detect]
def DetectOutline(self, image):
  image_size = cv.GetSize(image)
  grayscale = cv.CreateImage(image_size, 8, 1)	%\label{Detect:grayscale})
  cv.CvtColor(image, grayscale, cv.CV_BGR2GRAY)
  cv.EqualizeHist(grayscale, grayscale)	%\label{Detect:equalize})
  storage = cv.CreateMemStorage(0)
  cv.Threshold(grayscale, grayscale, 50, 255, cv.CV_THRESH_BINARY) %\label{Detect:threshold})
  self.contours = cv.FindContours(grayscale, %\label{Detect:contours})
    cv.CreateMemStorage(),
    cv.CV_RETR_TREE,
    cv.CV_CHAIN_APPROX_SIMPLE)
  if len(self.contours) > 0:
    self.contours = cv.ApproxPoly (self.contours, %\label{Detect:approx})
      storage,
      cv.CV_POLY_APPROX_DP,
      1.5,
      1)
  return self.contours
\end{lstlisting}

In function \lstinline{DetectOutline()} in Listing~\ref{Detect} is shown how to
find outlines of objects in image.
First we convert image to grey scale as seen on Listing~\ref{Detect},
line~\ref{Detect:grayscale}.

Then we run histogram equalization (Listing~\ref{Detect},
line~\ref{Detect:equalize}). Equalization makes objects better visible and gives
better output for thresholding (Listing~\ref{Detect},
line:~\ref{Detect:threshold}) which makes black and white as you can see in
Fig.~\ref{fig:histogram}.

Outline detection is done with function
\lstinline{cv.FindContours()} (Listing~\ref{Detect},
line:~\ref{Detect:contours}). Output of thresholding is shown in
Fig.~\ref{fig:threshold}.

\begin{figure}[h]
  \center
  \subfloat[Equalize histogram]{
    \label{fig:histogram}
    \includegraphics[width=0.48\columnwidth]{images/3test-equalize-histogram.png}
  }
  \subfloat[Thresholding]{
    \label{fig:threshold}
    \includegraphics[width=0.48\columnwidth]{images/4test-threshold.png}
   }
\caption{Effects applied on images}
\label{fig:operations}
\end{figure}

After outline detection we have a tree of contours stored in the variable
\lstinline{self.contours}. These trees are iterable objects sorted from outer to
inner outline connected by property h\_next and v\_next that we will describe in
paragraph about creation of objects from outlines.

Contour can be very complicated and consist of thousands of points, which
would cause objets with thousands of vertices. It is time demanding to
simulate complicated objects, that is why we use polynomial approximation of
the contour points. (line:~\ref{Detect:approx}). Visualisation of outlines is
shown in Fig.~\ref{fig:contours}.

\begin{figure}[h]
\center
\includegraphics[width=0.8\columnwidth]{images/5test-contours.png}
\caption{Visualisation of contours}
\label{fig:contours}
\end{figure}

Now we have all outlines stored in the outline tree structure, so we can create
objects and apply a physics.

\section{Physics simulation in Box2D}
There is lot of physics engines that can be used for simulation of physics.
Because we wanted to simulate physics only in 2D we could code our own
implementation of physics, or use one of commercial or open source engines. We
chose Box2D [\cite{GameEngines}], which is open source 2D physics engine with
implementation of rigid body objects and their collisions.
\subsection{World}
To create physics simulation we need to create world. World is object that
manages memory, objects and simulation. Creation of world is shown in Listing
\ref{World}:
\begin{lstlisting}[caption=Creation of Box2D world,label=World]
self.worldAABB=box2d.b2AABB()
self.worldAABB.lowerBound = (-100.0, -100.0) %\label{World:lower})
self.worldAABB.upperBound = ( 600.0, 600.0)	%\label{World:upper})
gravity = (0.0, -10.0) %\label{World:gravity})

doSleep = True %\label{World:sleep})
self.world = box2d.b2World(self.worldAABB, gravity, doSleep)
\end{lstlisting}
First we have to create boundaries of the world. We define them as vectors from
bottom left (Listing~\ref{World}, line:~\ref{World:lower}) to top right
(Listing~\ref{World}, line:~\ref{World:upper}). Objects have to be inside the
boundaries, when an object touch the boundary it gets stuck. Then we define
gravity vector (Listing~\ref{World}, line:~\ref{World:gravity}). The last
thing before creation of the world we allow objects to
sleep (Listing~\ref{World}, line:~\ref{World:sleep}). Object that are not moving
fall asleep, then they are ignored by the engine. Last line of
Listing~\ref{World} creates the world.

World is created and we are ready to create objects from outlines.

\subsection{Objects}
Every object that is simulated in Box2D consists of body and shapes. Our objects
are described by the contour tree. To create objects we need to iterate
through contour tree to find contours that belongs together and create
objects for these contours.

\subsection{Contour tree}
Contour tree is object in which are stored points of each contour. Contours are
connected by functions returning reference to other contours with
\lstinline{h_next()} and \lstinline{v_next()}, where \lstinline{h_next()} is
reference to deeper contour, and \lstinline{v_next()} is reference to another
object contour. To iterate over all contours we have created recursive function
shown in Listing: \ref{CreateObjects}.

\begin{lstlisting}[label=CreateObjects,caption=Function to iterate through contour tree]
def CreateObjectsFromCountours(self, cont, h=0, v=0):
  if v>0:	%\label{CreateObjects:depth})
    density = 10.0 %\label{CreateObjects:density10})
  else:
    density = 0 %\label{CreateObjects:density0})
  if len(cont)>8:
    self.CreateObject(cont,h,v)

  if cont.v_next():
    v += 1
    self.CreateObjectsFromCountours( cont.v_next(), h, v)
    v -= 1

  if cont.h_next():
    h += 1
    self.CreateObjectsFromCountours( cont.h_next(), h, v)
\end{lstlisting}

We iterate through contour tree. First level is outer contour of the image
(Listing \ref{CreateObjects} line:~\ref{CreateObjects:depth}), because we do not
want the outer contour to move, we set it as static by setting density to 0
(Listing \ref{CreateObjects} line:~ \ref{CreateObjects:density0}). Every other
contour is dynamic body with density set to 10 (Listing \ref{CreateObjects}
line:~\ref{CreateObjects:density10}).

When we know what type of object we will create, we can create bodies and
shapes for our contours.
\subsection{Bodies, shapes and collisions}
Bodies are backbone used by shapes. One body can contain more shapes, but one
shape could be attached to only one body. Box2d is rigid body physics engine,
that mean that shapes attached to body can not move against other, or body. Body
have position and velocity. Forces, torques and impulses can be applied to body
[\cite{Box2DManual}]. Bodies just hold the shapes and shapes are elements that
collide together.

Listings \ref{Objects1},\ref{Objects2},\ref{Objects3} shows process of object
creation:
\begin{lstlisting}[name=Objects,label=Objects1,caption=Creation of object from outer contour]
def CreateObject(self, cont, h,v):
  contM = []
  for point in cont:
    x = point[0]/30.0	%\label{Objects:scaling})
    y = point[1]/30.0
    contM.append((x,y))

  bd=box2d.b2BodyDef() %\label{Objects:CreateBody})
  bd.position = ( 0.0, 0.0 )

  edgeDef=box2d.b2EdgeChainDef() %\label{Objects:CreateChain})
  edgeDef.setVertices(contM) %\label{Objects:AssignVertices})

  if v==0:
    body = self.world.CreateBody(bd)
    try:
      self.contourBodies.append(body)
    except:
      self.contourBodies = [body] %\label{Objects:CountourBodies})
      body.CreateShape(edgeDef) %\label{Objects:AttachShape})
\end{lstlisting}
Image size is measured in pixels and Box2D units are kilograms, meters,
and seconds (KMS) we should scale images coordinates to fit in 0.1m to
10m. In that scale is performance of Box2D the best. We are doing it by
dividing of value of pixel coordinates by 30.0
(Listing~\ref{Objects1} line:~\ref{Objects:scaling})

Then we create a body definition that will represent our 
contour(Listing~\ref{Objects1} line:~ \ref{Objects:CreateBody}) and set up it
initial position in next line.

After that we create shape of body as chain of edges (Listing~\ref{Objects1}
line:~ \ref{Objects:CreateChain}) and assign the array of vertices to it
(Listing~\ref{Objects1} line:~ \ref{Objects:AssignVertices}). Edges are special
type of shapes that have no mass. Edges are represented as lines between
vertices that collide with other non-edge objects. Edges are easy to create
because they do not have to be concave unlike polygons.

At last we attach this shape to created body Listing~\ref{Objects1}
\ref{Objects:AttachShape}. Because Box2D does not keep track about body
definitions, we have to store bodies in to array for later use
(Listing~\ref{Objects1} line: \ref{Objects:CountourBodies}).

Listing of the function \lstinline{CreateObject()} continues in
Listing~\ref{Objects2}. This part of function creates dynamic objects inside the
outer contour. In this part we prepare list for bodies of objects, so we can
modify objects that are already created or objects that we want append new shapes.

\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects2,caption=Creation of objects]
if v == 1:
    try:
      body = self.objectBodies[h]
    except:
      body = self.world.CreateBody(bd)
      self.objectBodies[h] = body
\end{lstlisting}

\subsection{tessellation}
Box2D supports only collisions between convex objects and contours of objects
captured by camera are mostly not convex. So we have to break outlines to
convex polygons. There is more ways how to break concave objects. We chose the
2D constrained Delaunay triangulation algorithm implemented by poly2tri Python
library[\cite{Delauanay}]. Function CreateObject() continuous in
Listing~\ref{Objects3}

\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects3,caption=Creation of objects]
    polyline = []
    for (x,y) in cont:
      polyline.append(p2t.Point(x,y)) %\label{Objects:polyline})
    cdt = p2t.CDT(polyline) %\label{Objects:p2t})
    triangles = cdt.triangulate()
    for t in triangles:  %\label{Objects:scale1})
      x1 = t.a.x/30.0
      y1 = t.a.y/30.0
      x2 = t.b.x/30.0
      y2 = t.b.y/30.0
      x3 = t.c.x/30.0
      y3 = t.c.y/30.0  %\label{Objects:scale2})
      if math.hypot(x2-x1,y2-y1)<0.1: %\label{Objects:minSize1})
        x2 = x2 + math.copysign(0.1, x2-x1)
        y2 = y2 + math.copysign(0.1, y2-y1)
      if math.hypot(x3-x2,y3-y2)<0.1:
        x3 = x3 + math.copysign(0.1, x3-x2)
        y3 = y3 + math.copysign(0.1, y3-y2)
      if math.hypot(x1-x3,y1-y3)<0.1:
        x1 = x1 + math.copysign(0.1, x1-x3)
        y1 = y1 + math.copysign(0.1, y1-y3) %\label{Objects:minSize2})
      poly=box2d.b2PolygonDef()  %\label{Objects:polygondef})
      poly.setVertices(((x1, y1), (x2, y2), (x3, y3)))
      poly.density = 1.0
      poly.restitution = 0.0
      poly.friction = 0.0
      body.CreateShape(poly)  %\label{Objects:createshape})
      body.SetMassFromShapes() %\label{Objects:setMass})
\end{lstlisting}

The creation of objects continues with tessellation. We need to assign vertices
to structure that could be understood by poly2tri library
(Listing~\ref{Objects3} line:~\ref{Objects:polyline}) and we initialize the CDT
object (Listing~\ref{Objects3} line:~\ref{Objects:p2t}). In next line we call
function that will create triangles from the vertices assigned before. These
triangles are in image pixel coordinates, so we need to scale them at first
(Listing~\ref{Objects3} lines:~\ref{Objects:scale1}-\ref{Objects:scale2}). Now
when we have triangles scaled we need to scale the triangles that are too small
to triangles with size at least 0.1m because of speed optimalization, this is done in
Listing~\ref{Objects3} lines:~\ref{Objects:minSize1}-\ref{Objects:minSize2}).
Now we have set of triangular shapes that could be attached to body
(Listing~\ref{Objects3} line:~\ref{Objects:createshape}). Because these objects
are compound objects, we need to set the center and amount of mass to this 
body. We can let Box2D set this properties based on shape information with
function \lstinline{SetMassFromShapes()} (Listing~\ref{Objects3}
line:~\ref{Objects:setMass}). Visualisation of objects is in Fig.~\ref{fig:objects}.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{images/6dynamic-objets.png}
\caption{Visualisation of objects after triangulation}
\label{fig:objects}
\end{figure}

After this we can add other objects and start simulation by function \lstinline{Step()}.
After few second of simulation are all objects on the bottom of the screen like
is shown in Fig~\ref{fig:simulation}.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{images/7dynamic-objets-simulation.png}
\caption{Visualisation of simulation}
\label{fig:simulation}
\end{figure}

\section{Future work}
This approach can be used in interactive blackboards used for education of
physics on elementary schools, in future work we plan to implement
identification of some special objects like springs or joints.

We are planning to implement object tracking, and dynamic object morphing so we
could interact with simulated objects. Because of Box2D is rigid body engine,
is complicated to simulate physics of objects that change their shape in time.
We plan to implement some soft body elements to make this possible.

The last stage will be usage of captured images as source of textures of
simulated objects. Because this is only decorative element, we are planning to
implement this task as a last one.

\section*{Acknowledgments}
The work was supported by a grant (No. NIL-I-007-d) from Iceland, Liechtenstein 
and Norway through the EEA Financial Mechanism and the Norwegian Financial 
Mechanism. This project is also co-financed from the state budget of the Slovak
Republic.

%If you use BibTeX comment the second line, if not comment the first line
\bibliography{pcbib}
%\input{paper.bbl}

\end{document}